---
title: Adverse Event Data Visualization for ClinicalTrials.Gov Data
output:
  html_document:
    df_print: paged
  pdf_document: default
geometry: margin=2cm
fontsize: 11pt
mainfont: Calibri Light
graphics: yes
editor_options: 
  chunk_output_type: inline
---

<style type="text/css">
  body{
    font-family: Arial, Helvetica, sans-serif;
}
</style>


```{r setup, include=FALSE}
# renv::dependencies()
# renv::init()

# tidyverse and data manipulation
library("knitr")
library("kableExtra")
library("readxl")
library("stringr")
library("tidyverse")
library("dplyr")
library("writexl")
library("openxlsx")
library(jsonlite)

# table package
library("tibble")
library("tidyr")
library("gtsummary")

# plot packages
library(fmsb)
library(ggradar)
library(ggpubr)
library(ctrdata)
```


# Downloading Data
```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}

# URL of Trial: https://www.clinicaltrials.gov/study/NCT03515837

url_ctgov <- "https://www.clinicaltrials.gov/study/NCT03515837"
q <- ctrGetQueryUrl(url = url_ctgov)

dbc <- nodbi::src_sqlite(dbname = "my_db", collection = "my_coll")

ctrLoadQueryIntoDb(queryterm = q, con = dbc)

ae_fields <- c("resultsSection.adverseEventsModule.eventGroups.title",
               "resultsSection.adverseEventsModule.eventGroups.id",
               "resultsSection.adverseEventsModule.otherEvents.organSystem",
               "resultsSection.adverseEventsModule.otherEvents.term",
               "resultsSection.adverseEventsModule.otherEvents.stats",
               "resultsSection.adverseEventsModule.seriousEvents.organSystem",
               "resultsSection.adverseEventsModule.seriousEvents.term",
               "resultsSection.adverseEventsModule.seriousEvents.stats")

# In case you want other fields
# dbFindFields(namepart = "adverse", sample = FALSE, con = dbc)

result <- dbGetFieldsIntoDf(
  fields = ae_fields,
  con = dbc)

# making arm and arm description dataset (to append later)
arm <- trimws(unlist(strsplit(result$resultsSection.adverseEventsModule.eventGroups.id, "/")))
arm_desc <- trimws(unlist(strsplit(result$resultsSection.adverseEventsModule.eventGroups.title, "/")))
arm_df <- data.frame(arm, arm_desc)

# -- unnesting/unlisting the data -- #
ae_clean <- list()
for (type in c("other", "serious")){

  # subset to other or serious
  df_subset <- result %>% 
      select(contains(type))
  
  # remove long names
  names(df_subset) <- sub(".*\\.", "", names(df_subset))
  
  # Seperate the variables that are listed as one long string
  organSystem <- df_subset$organSystem
  term <- df_subset$term

  # Unnest Stats
  df_unnest <- df_subset %>% 
    unnest(stats) %>%
    rename(arm = groupId, organSystemName = organSystem)
  
  organSystem_clean <- rep(trimws(unlist(strsplit(organSystem, "/"))), each = length(unique(df_unnest$arm)))
  term_clean <- rep(trimws(unlist(strsplit(term, "/"))), each = length(unique(df_unnest$arm)))

  ae_clean[[type]] <- df_unnest %>% 
    mutate(organSystemName = organSystem_clean,
           term = term_clean,
           event_type = type)
}

ae_ctgov_df <- bind_rows(ae_clean) %>%
  # adding arm description for labeling
  left_join(arm_df, by = "arm")
```

## Radar plot

### Prepping the Data
This is an example of what the data looks like that was downloaded from clinicaltrials.gov:
```{r}
ae_ctgov_df %>%
  kable()  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
    scroll_box(width = "950px", height = "410px")
```

Since we don't have patient level data to find out how many individuals were in each category, we will use the highest count per category's term as a way to display "at least this many people had an AE in this category":

```{r}
max_term_per_cat <- ae_ctgov_df %>%
  group_by(event_type, organSystemName, arm) %>% 
  arrange(desc(numAffected), .by_group=TRUE) %>%
  filter(row_number()==1) %>%
  ungroup()
```

Now we are ready to begin building the dataset. Let's start by renaming the categories to something shorter, but still informative:
```{r}
# Using manual renaming from excel sheet
renaming_dat <- read_xlsx("Conversions//MEDRA_Word_Conversions.xlsx")

filtered_data <- max_term_per_cat %>%
  left_join(renaming_dat, by = "organSystemName") %>%
  select(-organSystemName) %>%
  rename(organSystemName = 'New Name') %>%
  relocate(organSystemName, .after=term) %>%
  arrange(desc(numAffected)) %>%
  
  # make pct var
  mutate(pct_affected = numAffected/numAtRisk) %>%   
  
  # this arm is experimental so removing for visualization purposes
  filter(!arm == "EG002")
```

Then we want to write a loop that takes in the data per event type, takes the ten most prelevant categories, makes each category a column with its own count, and creates a row per arm.
```{r}
event_data <- list()
spider_plot_dat_list <- list()

for (event in sort(unique(filtered_data$event_type))) {

# Filtering grade and events
filtered_event_dat <- filtered_data %>%
  filter(event_type == event) %>% 
  arrange(desc(numAffected))

# Taking the 10 most prevalent overall categories
categories <- filtered_event_dat %>% 
  group_by(organSystemName) %>% 
  mutate(total = sum(numAffected))  %>%
  arrange(desc(total))  %>%
  filter(row_number()==1)  %>%
  ungroup() %>%
  filter(row_number()==1:10) %>%
  pull(organSystemName) 

filtered_event_dat <- filtered_event_dat %>% 
  filter(organSystemName %in% categories)


# Pivoting to Wider for easiest formatting
event_data[[event]] <- filtered_event_dat %>% 
  select(organSystemName, arm_desc, pct_affected) %>% 
  pivot_wider(id_cols = organSystemName,
              names_from = arm_desc,
              # names_prefix = "Grade ",
              values_from = pct_affected,
              values_fill = 0) %>% 
  complete(organSystemName =categories) %>%
  arrange(desc(`Placebo + Pemetrexed + Chemo`), desc(`Pembro + Pemetrexed + Chemo`)) %>% 
  mutate(across(everything(), ~ replace_na(., 0)))

# Making the spider plot data
spider_plot_dat <- event_data[[event]]   %>%
  rename(Category = organSystemName) 

spider_dat <- as.data.frame(t(spider_plot_dat))
names(spider_dat) <- lapply(spider_dat[1, ], as.character)
spider_dat <- spider_dat[-1,] %>%
  add_column(Name = rownames(spider_dat[-1,]), .before=1) %>%
   mutate(across(-Name, ~ as.numeric(.)))
rownames(spider_dat) <- NULL
# colnames(spider_dat) <- sapply(colnames(spider_dat), function(x) strsplit(x, " ")[[1]][1])

spider_plot_dat_list[[paste(event)]] <- spider_dat %>% 
  mutate(Name = factor(Name,
                           levels = c("Placebo + Pemetrexed + Chemo",
                                      "Pembro + Pemetrexed + Chemo",
                                      "Pembro Monotherapy following PD on Placebo + Pemetrexed + Chemo")))

}
```


Here's what the loop produces for "Other Events" for example:
```{r}
spider_plot_dat_list$`other` %>% 
 kable()  %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
    scroll_box(width = "950px", height = "150px")
```

Now, we are ready to create the plots!

### Creating Plot

For creating a plot where each grade is an entirely separate plot, we would loop through each grade as so:

```{r}
color <- c("#619CFF", "#00BA38")
title <- names(spider_plot_dat_list)
spider_plots <- list()

# Create the radar chart
for(i in names(spider_plot_dat_list)){

color_fix <- color
if (nrow(spider_plot_dat_list[[i]]) < length(color)) {
  color_fix <- color[(length(color) - nrow(spider_plot_dat_list[[i]]) + 1):length(color)]}

spider_plots[[i]] <- ggradar(spider_plot_dat_list[[i]],
        background.circle.colour = "white",
        legend.position = "bottom",
        font.radar = "serif",
        grid.min = 0,
        grid.mid = 0.5,
        grid.max = 1,
        axis.label.size = 3.5,
        grid.label.size = 4,
        legend.text.size = 10,
        grid.line.width = 0.5,
        group.colours = color_fix,
        group.point.size = 2.3,
        group.line.width = 1.2,
        # axis.label.offset = 1.2,
        gridline.label.offset = -0.005,
        centre.y = -0.2)  +
        # ggtitle(paste(i, "AEs"))  +
        theme(plot.title = element_text(size = 12,
                                        hjust = 0.5,
                                        face = "bold")) +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "lines")) + 
guides(color = guide_legend(nrow = 3))
}
```


Here is the "other" plot for example:
```{r results = 'asis', fig.width=11, fig.height=9}
spider_plots$other
```

Using ggarrange to put them all together
```{r results = 'asis', fig.width=11, fig.height=9}
radar_together <- ggarrange(plotlist = spider_plots, widths = c(1.7, 1.6), common.legend = TRUE,
          legend = "bottom")

radar_together
```

## Love Plots

If the number of patients for each arm are not comparable (as in one has a much higher count than another), then a love plot is a good option for displaying the data


The data does not really need to be altered too much, we just want to add a few variables that calculate percent of patients affect by each adverse event out of their arm
```{r echo=TRUE}
love_plot_dat <- ae_ctgov_df   %>%   
  filter(!arm == "EG002") %>%
  group_by(term, event_type) %>%
  # creating a percent variable to calculate num effect by term/event type (ignoring arm)
  mutate(total_pct_events = (sum(numAffected)/sum(numAtRisk))) %>%
  arrange(total_pct_events) %>%
  ungroup() %>%
  mutate(pct_affeced_by_arm = numAffected/numAtRisk)
```

### Creating Plot

We are going to make one plot for "other events" and one for "serious"
```{r}
# for other events
other_chart <- love_plot_dat %>%
  filter(event_type == "other") %>%
  mutate(term = forcats::fct_inorder(term)) %>%
  # picking events that happened over 5% of the time
ggplot(data = .,
       aes(x = pct_affeced_by_arm*100,
           y = term,
           group = arm_desc,
           color = arm_desc)) +
geom_point(size = 2) + 
xlab("Percent of Patients Affected") +
geom_line(orientation = "y", linewidth = 0.9)  + 
  theme_bw() +
   ggtitle("Other AEs") +
  scale_color_manual(values=c("#619CFF", "#00BA38")) +
  guides(colour=guide_legend(title="Treatment Arm"))

serious_chart <- love_plot_dat %>%
  filter(event_type == "serious") %>%
  mutate(term = forcats::fct_inorder(term)) %>%
  # picking events that happened over 5% of the time
  filter(total_pct_events > 0.0025) %>%
ggplot(data = .,
       aes(x = pct_affeced_by_arm,
           y = term,
           group = arm,
           color = arm,
           fill = arm)) +
geom_point(size = 2)  + 
xlab("Percent of Patients Affected") + 
 scale_color_manual(values=c("#619CFF", "#00BA38")) +
  geom_line(orientation = "y", linewidth = 0.9)  + 
  theme_bw() +
  ggtitle("Serious AES")
```

```{r results = 'asis', fig.width=10, fig.height=6,fig.fullwidth=TRUE }
ggarrange(other_chart, serious_chart,
          legend = "bottom", common.legend=TRUE,
          ncol = 2)
```

# Arranging together

```{r results = 'asis', fig.width=10, fig.height=6,fig.fullwidth=TRUE }
final_plot <- ggarrange(other_chart + ggtitle(""), 
                        spider_plots$other + ggtitle(""),
                        common.legend = TRUE, legend = "bottom",
          widths = c(1, 1.5))

final_plot
```


<!-- Code for turning it into a png: -->
<!-- ```{r eval = false, echo=TRUE, include=FALSE} -->
<!-- png(filename = "filepath", -->
<!--     width = 12, height = 7, units = "in", res = 500) -->

<!-- annotate_figure(final_plot,  -->
<!--                 top = text_grob("Phase III trial of Pembrolizumab + Chemotherapy in nonsquamous NSCLC (KEYNOTE-789): Adverse Events",  -->
<!--                face = "bold", size = 14)) -->
<!-- dev.off() -->
<!-- ``` -->
