---
output:
  html_document:
    df_print: paged
  pdf_document: default
geometry: margin=2cm
fontsize: 11pt
mainfont: Calibri Light
graphics: yes
editor_options: 
  chunk_output_type: inline
---

<style type="text/css">
  body{
    font-family: Arial, Helvetica, sans-serif;
}
</style>


```{r setup, include=FALSE}
# tidyverse and data manipulation
library("knitr")
library("kableExtra")
library("readxl")
library("stringr")
library("tidyverse")
library("dplyr")
library("writexl")
library("openxlsx")

# table package
library("tibble")
library("tidyr")
library("gtsummary")

# plot packages
library(fmsb)
library(ggradar)
library(ggpubr)

# knitr options
opts_chunk$set(echo = TRUE, include = TRUE,
               warning = FALSE, message = FALSE,
               comment = "  ", prompt = TRUE)

# format numbers
opts_knit$set(root.dir = "./") # Change working directory here
knit_hooks$set(
  inline = function(x){
    if(!is.numeric(x)){x} 
    else{prettyNum(x, big.mark = ",")}
  }
)

# formatting options
options(round = 4,
        digits = 3,
        stringsAsFactors = FALSE)
```

# Data Step

```{r echo=FALSE}
source("G:\\KocherginskyM_984\\Analysis\\Code\\ctgov_function_code_script_02_19_24.R")
```

## Prepping your data for spider plots
```{r echo=FALSE}
dir_ <- "G://01 NUProtocols//NU18H02//"
dat_dir_ <- paste0(dir_, "Analysis//Data//Update 02.2022//")

dat_resp <- read.csv(paste0(dat_dir_, "NU 18H02-Data-Responses-02_24_2022 09_58.csv"))
dat_demo <- read.csv(paste0(dat_dir_, "NU 18H02-Arms_12_18_2023.csv"))
dat_eval <- read_xlsx(paste0(dat_dir_, "NU18H02_evaluability_12_18_2023.xlsx"))

#Dataset with renamed categories
renaming_dat <- read_xlsx("CTCAE_Word_Conversions.xlsx")

all_ae_dat <- dat_resp %>%
  filter(form_title == "Adverse Events TAK-659 & R-CHOP CRF") %>%
    select(ptid, response_set_id,	Timepoint, Status, starts_with("ae")) %>%
    mutate(ae_st_date = mdy(ae_st_date))

dt_consent_data <- dat_resp %>%
  filter(form_title == "Consent Form v1") %>%
  select(ptid, dt_consent) %>%
  mutate(dt_consent = mdy(dt_consent)) %>%
  group_by(ptid) %>%
  arrange(dt_consent) %>%
  filter(row_number() == 1)

dat_ae_merged <- merge(all_ae_dat, dt_consent_data, by = "ptid")
dat_ae_merged <- merge(dat_ae_merged, dat_demo %>%
                         select(ptid, d_arm), by = "ptid")

tox_eval_ptid <- dat_eval$`CASE NUM`[dat_eval$`Evaluable for toxicity` == "yes"]
# # Before Processing:
# dim(dat_ae_merged)
# length(unique(dat_ae_merged$ptid))

# Processing
processed_ae_data <- pre_processing_ae_data(dat_ae_merged)
# dim(processed_ae_data)
# length(unique(processed_ae_data$ptid))


# getting max ae per term per person (related)
maxae_data_multiarm <- processed_ae_data %>%
    filter(ae_rel_int1 %in% c("Possible","Probable" ,"Definite", 0 , 1, 2)) %>%
    # grouping by category + term
    group_by(ptid, organSystemName, term, d_arm) %>%
    # summarise by max grade
    dplyr::summarise(max_aegrade = max(ae_grade)) %>%
    ungroup()


# processed_ae_data %>%
#   select(ptid, d_arm) %>%
#   write_csv(path = "G:\\KocherginskyM_984\\Analysis\\Data\\id_dat.csv")
# 
# processed_ae_data %>%
#   select(ptid, term, organSystemName, ae_grade, ae_rel_int1) %>%
#   write_csv(path = "G:\\KocherginskyM_984\\Analysis\\Data\\ae_dat.csv")
```

```{r}
maxae_data_multiarm2 <- maxae_data_multiarm %>%
  left_join(renaming_dat, by = "organSystemName") %>%
  select(-organSystemName) %>%
  rename(organSystemName = 'New Name')  # %>%
  # mutate(max_aegrade = if_else(max_aegrade %in% c(1,2),
  #                              "1/2",
  #                              as.character(max_aegrade)))

# making the presentation categories
regroup <- maxae_data_multiarm2 %>%
  mutate(organSystemName = case_when(
    term %in% c("Lymphocyte count decreased", "Lymphocyte count increased",
                     "Anemia", "Platelet count decreased", "Neutrophil count decreased",
                     "Febrile neutropenia") ~  "Hematologic", 
    term %in% c("Aspartate aminotransferase increased", 
  "Alanine aminotransferase increased", 
  "Alkaline phosphatase increased",
  "Blood lactate dehydrogenase increased", 
  "Blood bilirubin increased", "Serum amylase increased", 
  "Lipase increased", "GGT increased", "Hypocalcemia", "Electrocardiogram QT corrected interval prolonged",
  "Sepsis", "Thrush", "Infections and infestations - Other, specify", "Urinary tract infection", 
  "Skin infection", "Upper respiratory infection", "Lung infection", "Fever", "Facial pain" , "Fatigue",
  "Weight loss", "Anorexia", "Myalgia", "Hypertension", "Diarrhea", "Nausea", "Vomiting") ~ "Non-\nHematologic",
  TRUE ~ NA)) %>%
  filter(!is.na(organSystemName))

new_ae <- rbind(maxae_data_multiarm2, regroup)
```

```{r}
# Taking the 10 most prevalent overall categories
categories <- new_ae %>% 
  group_by(ptid, organSystemName) %>% 
  filter(row_number()==1) %>%
  group_by(organSystemName) %>% 
  summarise(N = n()) %>% 
  arrange(desc(N)) %>%
  dplyr::filter(row_number()==1:10) %>%
  pull(organSystemName)

filtered_data <- new_ae %>% 
  filter(organSystemName %in% categories)
```

```{r}
total_counts <- c(
  "Dose Level 1 (60mg)" = 3,
  "Dose Level 2 (80mg)" = 3,
  "Dose Level 3 (100mg)" = 6
)

col_order <- NULL
grade_data <- list()
spider_plot_dat_list <- list()

for (grade in sort(unique(filtered_data$max_aegrade))) {

# Filtering grade and events
filtered_grade_dat <- filtered_data %>%
  filter(max_aegrade == grade) %>% 
  arrange(desc(max_aegrade))

grade_data[[grade]] <- filtered_grade_dat %>% 
  group_by(ptid, organSystemName) %>% 
  filter(row_number()==1) %>%
  group_by(organSystemName, d_arm) %>%
  mutate(N = n()) %>%
  select(organSystemName, d_arm, max_aegrade, N) %>% 
  filter(row_number()==1) %>% 
  ungroup() %>% 
  pivot_wider(id_cols = organSystemName,
              names_from = d_arm,
              # names_prefix = "Grade ",
              values_from = N,
              values_fill = 0) %>% 
  complete(organSystemName =categories) %>%
  # arrange(desc(`Dose Level 3 (100mg)`)) %>%
  mutate(across(everything(), ~ replace_na(., 0))) 

# Divide each arm's count by the total per arm to get proportions
grade_data[[grade]] <- grade_data[[grade]] %>%
    mutate(across(-organSystemName, ~ . / total_counts[cur_column()]))

  
# Making the spider plot data
spider_plot_dat <- grade_data[[grade]]  %>%
  rename(Category = organSystemName) 

spider_dat <- as.data.frame(t(spider_plot_dat))
names(spider_dat) <- lapply(spider_dat[1, ], as.character)
spider_dat <- spider_dat[-1,] %>%
  add_column(Name = rownames(spider_dat[-1,]), .before=1) %>%
   mutate(across(-Name, ~ as.numeric(.)))
rownames(spider_dat) <- NULL
# colnames(spider_dat) <- sapply(colnames(spider_dat), function(x) strsplit(x, " ")[[1]][1])

# spider_dat2 <- spider_dat[, c(1, order(-colSums(spider_dat[, -1]), -as.numeric(spider_dat[3, -1])) + 1)]

 # Set the column order only once
  if (is.null(col_order)) {
    col_order <- c(1, order(-colSums(spider_dat[, -1])) + 1)
  }

# Apply the stored column order
  spider_dat2 <- spider_dat[, col_order]

  spider_plot_dat_list[[paste("Grade ", grade)]] <- spider_dat2
}
```

## Plot A - All Seperate Arms 

For creating a plot where each grade is an entirely separate plot, we would loop through each grade as so:

```{r}
color <- c("#F0E442", "#E69F00", "firebrick3")
title <- names(spider_plot_dat_list)
spider_plots <- list()

# Create the radar chart
for(i in names(spider_plot_dat_list)){

color_fix <- color
if (nrow(spider_plot_dat_list[[i]]) < length(color)) {
  color_fix <- color[(length(color) - nrow(spider_plot_dat_list[[i]]) + 1):length(color)]}

spider_plots[[i]] <- ggradar(spider_plot_dat_list[[i]],
        background.circle.colour = "white",
        legend.position = "bottom",
        font.radar = "serif",
        grid.min = 0,
        grid.mid = 0.5,
        grid.max = 1,
        axis.label.size = 12,
        grid.label.size = 12,
        legend.text.size = 12,
        grid.line.width = 1.5,
        group.colours = color_fix,
        group.point.size = 6.5,
        group.line.width = 4.5,
        # axis.label.offset = 1.2,
        gridline.label.offset = 0.2,
        centre.y = -0.25,
        point.alpha = 0.9,
        line.alpha = 0.8)  +
        ggtitle(paste(i))  + 
        theme(plot.title = element_text(size = 35,
                                        hjust = 0.5,
                                        face = "bold")) +
   theme(plot.margin = margin(0,0,0,0, 'cm'))
}
```


Using ggarrange to put them all together (plot_grid also works)
```{r}
allgrades_plot <- ggarrange(plotlist = spider_plots, widths = c(0.5,0.5), common.legend = TRUE,
          legend = "bottom")

plot_grid(
  ggarrange(plotlist = list(spider_plots[[1]], ggplot() + theme_void(), spider_plots[[2]]), 
            ncol=3, legend = "none",
            widths = c(1,0.01,1)),
  ggarrange(plotlist = list(spider_plots[[3]], NULL, spider_plots[[4]]), ncol=3, legend = "none"),
  ncol=2
  )

allgrades_plot <- ggarrange(plotlist = list(spider_plots[[1]], 
                                           ggplot() + theme_void(), 
                                           spider_plots[[2]],
                                           spider_plots[[3]], 
                                           ggplot() + theme_void(), 
                                           spider_plots[[4]]), 
                            widths = c(1, 0.001, 1, 1, 0.001, 1), 
                            common.legend = TRUE, legend = "bottom")


allgrades_plot
```


# Waterfall Plots

## Prepping the Data

The data does not really need to be altered too much, we just want to add a few variables that calculate percent of patients affect by each adverse event out of their arm
```{r echo=TRUE}
love_plot_dat <- maxae_data_multiarm2 %>%
  mutate(term = gsub(" - Other, specify", "", term),
         max_aegrade = paste("Grade", max_aegrade)) %>%
    # max_aegrade = case_when(
    #   max_aegrade %in% c(3, 4) ~ "3/4",
    #   TRUE ~ max_aegrade)) %>%
  group_by(organSystemName, term, d_arm, max_aegrade) %>%
  summarise(n_pt = length(unique(ptid)), .groups = "drop") %>% 
  # Complete missing combinations with 0
  mutate(arm_sum = case_when(
    d_arm == "Dose Level 1 (60mg)" ~ 3,
    d_arm == "Dose Level 2 (80mg)" ~ 3,
    d_arm == "Dose Level 3 (100mg)" ~ 6),
  pct_affeced_by_arm =  n_pt/arm_sum)  %>%   
  group_by(d_arm, term) %>%
  # creating a percent variable to calculate num effect by term/event type (ignoring arm)
  mutate(total_pct_events = sum(pct_affeced_by_arm)) %>%
  ungroup() %>% arrange(total_pct_events)
```

### Creating Plot

We are going to make one plot per arm
```{r}

love_plots <- list()
for (i in sort(unique(love_plot_dat$d_arm))){
  
love_plots[[i]] <- love_plot_dat %>%
  filter(d_arm == i) %>%
  group_by(term) %>% 
  mutate(pct_total = sum(n_pt)) %>%
  arrange(pct_total) %>%
  ungroup() %>%
  mutate(term = forcats::fct_inorder(term)) %>%
   # picking events that happened over 5% of the time
  filter(total_pct_events > 0.4) %>%
  complete(nesting(organSystemName, term), d_arm, max_aegrade, fill = list(pct_affeced_by_arm = 0)) %>%
ggplot(data = .,
       aes(x = pct_affeced_by_arm*100,
           y = term,
           group = max_aegrade,
           color = max_aegrade)) +
geom_point(size = 2) + 
expand_limits(x=c(0,100)) +
scale_x_continuous(breaks = c(0, 20, 40, 60 , 80, 100)) +
xlab("Percent of Patients Affected") +
geom_line(orientation = "y", linewidth = 0.9)  + 
  theme_bw() +
  ggtitle(paste(i)) +
  scale_color_manual(values=c("#7AD151FF", "#2A788EFF", "#440154FF")) +
  # scale_colour_viridis_d(direction = -1) +
  guides(colour=guide_legend(title="Grade")) +
  theme(text = element_text(size = 15))
}
```


```{r results = 'asis', fig.width=15, fig.height=10,fig.fullwidth=TRUE }
date <-  gsub("/", "", format(as.Date(Sys.Date()), "%m/%d/%y"))


love_plots_arrange <- ggarrange(plotlist=love_plots,
          legend = "none",
          ncol = 3)
love_plots_arrange

png(filename = file.path("G://KocherginskyM_984//Documentation//", paste0("love_plot_all_", date, ".png")),
    width = 25, height = 20, units = "in", res = 500)
print(love_plots_arrange)
dev.off()

png(filename = file.path("G://KocherginskyM_984//Documentation//", paste0("love_plot_dose3", date, ".png")),
    width = 10, height = 11, units = "in", res = 500)
print(love_plots[["Dose Level 3 (100mg)"]])
dev.off()

png(filename = "G://KocherginskyM_984//Documentation//love_plot_NU18H02_justone_042525.png",
    width = 10, height = 11, units = "in", res = 500)
print(love_plot_total)
dev.off()

png(filename = "G://KocherginskyM_984//Documentation//love_plot_NU18H02_grade3_042425.png",
    width = 10, height = 11, units = "in", res = 500)
print(love_plots[["3/4"]])
dev.off()

png(filename = "G://KocherginskyM_984//Documentation//love_plot_NU18H02_grade4_042425.png",
    width = 10, height = 11, units = "in", res = 500)
print(love_plots[["4"]])

dev.off()



png(filename = "G://KocherginskyM_984//Documentation//radar_plot_NU18H02_041625.png",
    width = 12, height = 7, units = "in", res = 500)
annotate_figure(allgrades_plot, top = text_grob("Treatment-related adverse events from the
combination of TAK-659 and R-CHOP at each dose level", 
               face = "bold", size = 14))
dev.off()
```


```{r results = 'asis', fig.width=15, fig.height=10,fig.fullwidth=TRUE }
for (i in names(spider_plots)){
png(filename = file.path("G://KocherginskyM_984//Documentation//", paste0("radarplot_", gsub("/", "", i), "_", date, ".png")),
    width = 16, height = 16, units = "in", res = 500)
print(spider_plots[[i]])
dev.off()
}

for (i in names(spider_plots_combined)){
png(filename = file.path("G://KocherginskyM_984//Documentation//", paste0("radarplot_", gsub("/", "", i), "_", date, ".png")),
    width = 16, height = 16, units = "in", res = 500)
print(spider_plots_combined[[i]])
dev.off()
}
```